Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-08 popl24/minicat.tex
Update minicat.tex

\subsection{Syntax} The syntax of $\minifed$, defined in
Figure \ref{fig-minifed-syntax}, includes a standard boolean algebra
with $\eand$, $\eor$, $\exor$, and $\enot$ as primitives. In addition,
programs have different kinds of variables including \emph{secrets}
$\secret{\cid}{w}$, \emph{flips} $\flip{\cid}{w}$, and \emph{views}
$\view{\cid}{w}$.  Each of these variables is indexed by an ``owner''
client $\cid$ and distinguishing string $w$. So for example,
$\secret{\ttt{1}}{\ttt{foo}}$ is client $\ttt{1}$'s secret called
$\ttt{foo}$. All clients can make reference to the \emph{oracle}
through variables $\oracle{w}$- this is a form of shared randomness
that is standard in the MPC setting \cite{evans2018pragmatic}.  Each client can only
compute on their own variables in \emph{expressions} $\be$, and share
values with other clients by assignment to their views as in
$\eassign{\view{\cid}{w}}{\be}$.  A \emph{protocol} $\prog$ is a
possibly empty sequence of view assignments. We will generally omit
$\varnothing$ from example code, writing $\instr_1;\ldots;\instr_n$
instead of $\instr_1;\ldots;\instr_n;\varnothing$, and abusing notation we will
write $\prog_1;\prog_2$ to denote the concatenation of $\prog_1$
and $\prog_2$.

@@

Here we introduce some examples to illustrate $\minifed$ and how it
can be used to model secure communication protocols for values in the
binary field. Most of these are translations (into $\minifed$) of
object examples in previous related work
\cite{barthe2019probabilistic,darais2019language}.  We will return to
these as running examples throughout the paper to illustrate various
concepts and advantages of our methods. In Section
\ref{section-metalang} we will formulate a larger example
implementation of Yao's Garbled Circuits.
\begin{example}[One-Time Pad]
  \label{example-otp}
As observed in \cite{barthe2019probabilistic} and elsewhere, coin flips with $\exor$ can
be used to model one-time pad encryption in the binary field. On the
basis of this we can model secure communication between over an
insecure channel using symmetric key encryption. In the following
program, we assume $H = \{ 1,2 \}$ and $C = \{ 0 \}$. Client 1 first
shares the randomly generated ``key'' $\fx{1}{0}$ with client 2
through view $\view{2}{0}$. Then, client 1 encrypts its secret
$\sx{1}{0}$ and sends it to party 0, which relays the ciphertext in
$\vx{0}{0}$ to client 2. Finally, client 2 decrypts the ciphertext
and stores the plaintext in $\vx{2}{2}$.
\begin{verbatimtab}
v[2,"0"] := f[1,"0"];
v[0,"0"] := f[1,"0"] xor s[1,"0"];
v[2,"1"] := v[0,"0"];
v[2,"2"] := v[2,"1"] xor v[2,"0"]
\end{verbatimtab}
Intuitively, this protocol is \emph{correct} because $\vx{2}{2}$ is
exactly correlated with the secret $\sx{1}{0}$, and it is
\emph{secure} because the corrupt view $\vx{0}{0}$ in isolation is
probabilistically independent of the secret $\sx{1}{0}$.
\end{example}

\begin{example}[$\lambda_{\mathrm{obliv}}$]
  \label{example-lambda-obliv}
  In \cite{darais2019language}, Figure 3, two examples are introduced that illustrate
  subtleties of probabilistic programming in
  $\lambda_{\mathrm{obliv}}$, a language for oblivious memory
  management. We consider their translations here where we
  assume $H = \{ 1 \}$ and $C = \{ 0 \}$. First, the following
  example shows that the reuse of coin flips can result in the release of
  information, even though flips are chosen from a uniform
  distribution. This is because the adversary can deduce that
  $\secret{1}{s}$ is more likely to be 1 in case their views are the
  same value. We call this subexample $(a)$:
  \begin{verbatimtab}
    v[0,"0"] := select(s[1,"s"], flip[1,"sx"], flip[1,"sy"]);
    v[0,"1"] := flip[1,"sx"] \end{verbatimtab}
  In the next example, since the nested select is more likely to evaluate
  to $1$ than $0$, the adversary can deduce that $\secret{1}{s}$ is
  more likely to be 1 in case $\view{0}{0}$ is 1. We call this subexample $(b)$:
  \begin{verbatimtab}
  v[0,"0"] := select(s[1,"s"],select(flip[1,"sx"], flip[1,"sx"], flip[1,"sy"]),flip[1,"z"]) \end{verbatimtab}
  We note that these examples are considered ``bad'' in \cite{darais2019language} since there
  no information about honest secrets should be revealed to the adversary in
  their threat model. For this reason, they disallow the reuse of flips through a
  linear type discipline, which is a culprit in both of these examples. However,
  as we have shown in Example \ref{example-otp} the reuse of flips is useful in
  the modeling of symmetric key encryption, and generally we will not enforce
  the same linear type discipline on the use of flips.
\end{example}

\begin{example}[Oblivious Transfer]
  \label{example-OT}
  In \cite{barthe2019probabilistic} a protocol for $\ttt{OT}$ is
  defined that we recreate as follows, where client 2 is the receiver
  and client 1 is the sender, and a trusted third party provides the
  sender and receiver with shared randomness. The secret \ttt{s[2,c]}
  is the receiver's choice bit, and \ttt{s[1,m0]} and \ttt{s[1,m1]}
  are the sender's two message options that the receiver chooses from. 
\begin{verbatimtab}
  v[1,"r0"] := s[3,"r0"];
  v[1,"r1"] := s[3,"r1"];
  v[2,"d"] := flip[3,"d"];
  v[2,"rd"] := select(flip[3,"d"], s[3,"r1"], s[3,"r0"]);
  v[1,"e"] := s[2,"c"] xor v[2,"d"];
  v[2,"f0"] := s[1,"m0"] xor select(v[1,"e"],v[1,"r1"],v[1,"r0"]);
  v[2,"f1"] := s[1,"m1"] xor select(not v[1,"e"],v[1,"r1"],v[1,"r0"]);
  v[2,"mc"] := select(s[2,"c"],v[2,"f1"],v[2,"f0"]] xor v[2,"rd")
\end{verbatimtab}
Below we will discuss how to automatically verify security properties
of this protocol that correspond to the properties proven manually in
\cite{barthe2019probabilistic}, and additionally how to automatically
verify its passive security in the simulator model. With respect to
our language primitives, we note that this protocol is equivalent to
the following:
\begin{verbatimtab}  
  v[2,"mc"] := OT(s[2,"c"], s[1,"m1"], s[1,"m0"])
\end{verbatimtab}
\end{example}

>>>>

\subsection{Syntax} The syntax of $\minifed$, defined in
Figure \ref{fig-minifed-syntax}, includes a standard boolean algebra
with $\eand$, $\eor$, $\exor$, and $\enot$ as primitives. In addition,
programs have different kinds of variables including \emph{secrets}
$\secret{\cid}{w}$, \emph{flips} $\flip{\cid}{w}$, and \emph{views}
$\view{\cid}{w}$.  Each of these variables is indexed by an ``owner''
client $\cid$ and distinguishing string $w$. So for example,
$\secret{\ttt{1}}{\ttt{"foo"}}$ is client $\ttt{1}$'s secret 
$\ttt{"foo"}$. All clients can make reference to the \emph{oracle}
through variables $\oracle{w}$- this is a form of shared randomness
that is standard in the MPC setting \cite{evans2018pragmatic}.  Each client can only
compute on their own variables in \emph{expressions} $\be$, and share
values with other clients by assignment to their views as in
$\eassign{\view{\cid}{w}}{\be}$.  A \emph{protocol} $\prog$ is a
possibly empty sequence of view assignments. We will generally omit
$\varnothing$ from example code, writing $\instr_1;\ldots;\instr_n$
instead of $\instr_1;\ldots;\instr_n;\varnothing$, and abusing notation we will
write $\prog_1;\prog_2$ to denote the concatenation of $\prog_1$
and $\prog_2$.

@@

Here we introduce some examples to illustrate $\minifed$ and how it
can be used to model secure communication protocols for values in the
binary field. Most of these are translations (into $\minifed$) of
object examples in previous related work
\cite{barthe2019probabilistic,darais2019language}.  We will return to
these as running examples throughout the paper to illustrate various
concepts and advantages of our methods. In Section
\ref{section-metalang} we will formulate a larger example
implementation of Yao's Garbled Circuits.
\begin{example}[One-Time Pad]
  \label{example-otp}
As observed in \cite{barthe2019probabilistic} and elsewhere, coin flips with $\exor$ can
be used to model one-time pad encryption in the binary field. On the
basis of this we can model secure communication between over an
insecure channel using symmetric key encryption. In the following
program, we assume $H = \{ 1,2 \}$ and $C = \{ 0 \}$. Client 1 first
shares the randomly generated ``key'' $\fx{1}{0}$ with client 2
through view $\vx{2}{0}$. Then, client 1 encrypts its secret
$\sx{1}{0}$ and sends it to party 0, which relays the ciphertext in
$\vx{0}{0}$ to client 2. Finally, client 2 decrypts the ciphertext
and stores the plaintext in $\vx{2}{2}$.
\begin{verbatimtab}
v[2,"0"] := f[1,"0"];
v[0,"0"] := f[1,"0"] xor s[1,"0"];
v[2,"1"] := v[0,"0"];
v[2,"2"] := v[2,"1"] xor v[2,"0"]
\end{verbatimtab}
Intuitively, this protocol is \emph{correct} because $\vx{2}{2}$ is
exactly correlated with the secret $\sx{1}{0}$, and it is
\emph{secure} because the corrupt view $\vx{0}{0}$ in isolation is
probabilistically independent of the secret $\sx{1}{0}$.
\end{example}

\begin{example}[$\lambda_{\mathrm{obliv}}$]
  \label{example-lambda-obliv}
  In \cite{darais2019language}, Figure 3, two examples are introduced that illustrate
  subtleties of probabilistic programming in
  $\lambda_{\mathrm{obliv}}$, a language for oblivious memory
  management. We consider their translations here where we
  assume $H = \{ 1 \}$ and $C = \{ 0 \}$. First, the following
  example shows that the reuse of coin flips can result in the release of
  information, even though flips are chosen from a uniform
  distribution. This is because the adversary can deduce that
  $\sx{1}{s}$ is more likely to be 1 in case their views are the
  same value. We call this subexample $(a)$:
  \begin{verbatimtab}
    v[0,"0"] := select(s[1,"s"], flip[1,"sx"], flip[1,"sy"]);
    v[0,"1"] := flip[1,"sx"] \end{verbatimtab}
  In the next example, since the nested select is more likely to evaluate
  to $1$ than $0$, the adversary can deduce that $\sx{1}{s}$ is
  more likely to be 1 in case $\vx{0}{0}$ is 1. We call this subexample $(b)$:
  \begin{verbatimtab}
  v[0,"0"] := select(s[1,"s"], 
  select(flip[1,"sx"], flip[1,"sx"], flip[1,"sy"]),flip[1,"z"]) \end{verbatimtab}
  We note that these examples are considered ``bad'' in \cite{darais2019language} since there
  no information about honest secrets should be revealed to the adversary in
  their threat model. For this reason, they disallow the reuse of flips through a
  linear type discipline, which is a culprit in both of these examples. However,
  as we have shown in Example \ref{example-otp} the reuse of flips is useful in
  the modeling of symmetric key encryption, and generally we will not enforce
  the same linear type discipline on the use of flips.
\end{example}

\begin{example}[Oblivious Transfer]
  \label{example-OT}
  In \cite{barthe2019probabilistic} a protocol for $\ttt{OT}$ is
  defined that we recreate as follows, where client 2 is the receiver
  and client 1 is the sender, and a trusted third party provides the
  sender and receiver with shared randomness. The secret $\sx{2}{c}$
  is the receiver's choice bit, and $\sx{1}{m0}$ and $\sx{1}{m1}$
  are the sender's two message options that the receiver chooses from. 
\begin{verbatimtab}
  v[1,"r0"] := s[3,"r0"];
  v[1,"r1"] := s[3,"r1"];
  v[2,"d"] := flip[3,"d"];
  v[2,"rd"] := select(flip[3,"d"], s[3,"r1"], s[3,"r0"]);
  v[1,"e"] := s[2,"c"] xor v[2,"d"];
  v[2,"f0"] := s[1,"m0"] xor select(v[1,"e"],v[1,"r1"],v[1,"r0"]);
  v[2,"f1"] := s[1,"m1"] xor select(not v[1,"e"],v[1,"r1"],v[1,"r0"]);
  v[2,"mc"] := select(s[2,"c"],v[2,"f1"],v[2,"f0"]] xor v[2,"rd")
\end{verbatimtab}
Below we will discuss how to automatically verify security properties
of this protocol that correspond to the properties proven manually in
\cite{barthe2019probabilistic}, and additionally how to automatically
verify its passive security in the simulator model. With respect to
our language primitives, we note that this protocol is equivalent to
the following:
\begin{verbatimtab}  
  v[2,"mc"] := OT(s[2,"c"], s[1,"m1"], s[1,"m0"])
\end{verbatimtab}
\end{example}
